---
title: "AGV_AQ_Volame3"
output: html_notebook
---

```{r}
#I- Lirbrary 
library(ggplot2)
library(ggpubr)
library(readxl)
library(gtsummary)
library(tidyverse)
library(ggpubr)
library(rstatix)

#II- Lecture des données brutes
volameAGV <- read_excel("AGV_DOA.xlsx")
View(volameAGV)
```


```{r}
#Boxplot des proportion des AGV 

volameAGV$Time<-as.character(volameAGV$Time)

Totaux_t <- ggplot(volameAGV, aes(x=factor(Time, level=c('3', '10', '13')), y= `Total VFA (  )`, fill=Groups)) + 
    geom_boxplot() + theme_classic() +xlab('Weeks of age')

Pro_t <- ggplot(volameAGV, aes(x=factor(Time, level=c('3', '10', '13')), y= `Propionate (%)`, fill=Groups)) + 
    geom_boxplot() + theme_classic() +xlab('Weeks of age')

Ac_t <- ggplot(volameAGV, aes(x=factor(Time, level=c('3', '10', '13')), y= `Acetate (%)`, fill=Groups)) + 
    geom_boxplot()+ theme_classic() +xlab('Weeks of age')

Buty_t <- ggplot(volameAGV, aes(x=factor(Time, level=c('3', '10', '13')), y=`Butyrate (%)` , fill=Groups)) + 
    geom_boxplot() +theme_classic() +xlab('Weeks of age')

isoButy_t <- ggplot(volameAGV, aes(x=factor(Time, level=c('3', '10', '13')), y= `Isobutyrate (%)` , fill=Groups)) + 
    geom_boxplot() +theme_classic() +xlab('Weeks of age')

isoval_t <- ggplot(volameAGV, aes(x=factor(Time, level=c('3', '10', '13')), y= `Isovalerate (%)`, fill=Groups)) + 
    geom_boxplot() +theme_classic() + xlab('Weeks of age')


val_t <- ggplot(volameAGV, aes(x=factor(Time, level=c('3', '10', '13')), y= `Valerate (%)`, fill=Groups)) + 
    geom_boxplot()+  theme_classic() +xlab('Weeks of age')

capro_t <- ggplot(volameAGV, aes(x=factor(Time, level=c('3', '10', '13')), y= `Caproate (%)`, fill=Groups)) + 
    geom_boxplot()+  theme_classic() +xlab('Weeks of age')


cowplot::plot_grid(Totaux_t, Ac_t,Pro_t ,Buty_t,isoButy_t,val_t,isoval_t,capro_t, nrow = 4, ncol = 2, scale = .8)

```



```{r}
#Table moyenne des donnée + SD
sumVFAtoto<- volameAGV %>%
  group_by(Time, Groups) %>%
  get_summary_stats(`Total VFA (  )` ,`Acetate (%)`,`Butyrate (%)`,`Propionate (%)`,`Isovalerate (%)`, `Caproate (%)`,`Isovalerate (%)`,`Isobutyrate (%)`, type = "mean_sd")

sumVFAtoto
```



The repeated-measures ANOVA is used for analyzing data where same subjects are measured more than once. This test is also referred to as a within-subjects ANOVA or ANOVA with repeated measures. The “within-subjects” term means that the same individuals are measured on the same outcome variable under different time points or conditions.




Assumption
```{r}
colnames(volameAGV)[9] ="VFA_t"
colnames(volameAGV)[10] ="VFA_Acetate"
colnames(volameAGV)[11] ="VFA_propionate"
colnames(volameAGV)[12] ="VFA_isobutyrate"
colnames(volameAGV)[13] ="VFA_butyrate"
colnames(volameAGV)[14] ="VFA_isovalerate"
colnames(volameAGV)[15] ="VFA_valerate"
colnames(volameAGV)[16] ="VFA_caproate"

#Outliers
volameAGV %>%
  group_by(Time, Groups) %>%
  identify_outliers(VFA_caproate)

# Normality 
ggplot (data = volameAGV, aes (sample = VFA_caproate )) +
  stat_qq ()
volameAGV %>%
  group_by(Groups, Time) %>%
  shapiro_test(VFA_caproate)

##Createqqplot for each cell of design
ggqqplot(volameAGV, "VFA_caproate", ggtheme = theme_bw()) +
  facet_grid(Time ~ Groups, labeller = "label_both")
```
The score was normally distributed at each time point (p > 0.05), except for ctr treatment at t1, as assessed by Shapiro-Wilk’s test+ Plus la courbe est rectiligne, plus cela signifie que les quantiles de la variable sont similaires à ceux de la distribution attendue. 

VFAtotaux :ok 

From the plot above, as all the points fall approximately along the reference line, we can assume normality.
```{r}
###Kurskall wallis test between groups
kruskal.test(VFA_t ~ Time, data = volameAGV)
kruskal.test(VFA_Acetate ~ Time, data = volameAGV)   
kruskal.test(VFA_propionate ~ Time, data = volameAGV)
kruskal.test(VFA_isobutyrate ~ Time, data = volameAGV)
kruskal.test(VFA_butyrate ~ Time, data = volameAGV)
kruskal.test(VFA_isovalerate ~ Time, data = volameAGV)   
kruskal.test(VFA_valerate ~ Time, data = volameAGV)
kruskal.test(VFA_caproate ~ Time, data = volameAGV)

kruskal.test(VFA_t ~ Groups, data = volameAGV)
kruskal.test(VFA_Acetate ~ Groups, data = volameAGV)   
kruskal.test(VFA_propionate ~ Groups, data = volameAGV)
kruskal.test(VFA_isobutyrate ~ Groups, data = volameAGV)
kruskal.test(VFA_butyrate ~ Groups, data = volameAGV)
kruskal.test(VFA_isovalerate ~ Groups, data = volameAGV)   
kruskal.test(VFA_valerate ~ Groups, data = volameAGV)
kruskal.test(VFA_caproate ~ Groups, data = volameAGV)

kruskal.test(VFA_t ~ Lot_x_Time, data = volameAGV)
kruskal.test(VFA_Acetate ~ Lot_x_Time, data = volameAGV)   
kruskal.test(VFA_propionate ~ Lot_x_Time, data = volameAGV)
kruskal.test(VFA_isobutyrate ~ Lot_x_Time, data = volameAGV)
kruskal.test(VFA_butyrate ~ Lot_x_Time, data = volameAGV)
kruskal.test(VFA_isovalerate ~ Lot_x_Time, data = volameAGV)   
kruskal.test(VFA_valerate ~ Lot_x_Time, data = volameAGV)
kruskal.test(VFA_caproate ~ Lot_x_Time, data = volameAGV)


#From the output of the Kruskal-Wallis test, we know that there is a significant difference between groups, but we don’t know which pairs of groups are different. A significant Kruskal-Wallis test is generally followed up by Dunn’s test to identify which groups are different. It’s also possible to use the Wilcoxon’s test to calculate pairwise comparisons between group levels with corrections for multiple testing.

###Pairwise
pairwise.wilcox.test(volameAGV$VFA_t, volameAGV$Lot_x_Time,p.adjust.method = "BH")
pairwise.wilcox.test(volameAGV$VFA_Acetate, volameAGV$Lot_x_Time,p.adjust.method = "BH")
pairwise.wilcox.test(volameAGV$VFA_propionate , volameAGV$Lot_x_Time,p.adjust.method = "BH")
pairwise.wilcox.test(volameAGV$VFA_isobutyrate, volameAGV$Lot_x_Time,p.adjust.method = "BH")
pairwise.wilcox.test(volameAGV$VFA_butyrate, volameAGV$Lot_x_Time,p.adjust.method = "BH")
pairwise.wilcox.test(volameAGV$VFA_isovalerate, volameAGV$Lot_x_Time,p.adjust.method = "BH")
pairwise.wilcox.test(volameAGV$VFA_valerate , volameAGV$Lot_x_Time,p.adjust.method = "BH")
pairwise.wilcox.test(volameAGV$VFA_caproate, volameAGV$Lot_x_Time,p.adjust.method = "BH")
```
 
Caproate/ valertae : pas ok sur normalité (qqplot + test), mais ggqqplot good 
```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)

res.aov_to <-volameAGV %>% anova_test(dv = VFA_totaux, wid = EDE, within = Time, between =groups)
get_anova_table(res.aov_to)

res.aov_to <-volameAGV %>% anova_test(dv = VFAt_Acetate, wid = EDE, within = Time, between =groups)
get_anova_table(res.aov_to)


res.aov_to <-volameAGV %>% anova_test(dv = VFAt_caproate, wid = EDE, within = Time, between =groups)
get_anova_table(res.aov_to)


res.aov_to <-volameAGV %>% anova_test(dv = VFAt_propionate, wid = EDE, within = Time, between =groups)
get_anova_table(res.aov_to)


res.aov_to <-volameAGV %>% anova_test(dv = VFAt_isobutyrate, wid = EDE, within = Time, between =groups)
get_anova_table(res.aov_to)


res.aov_to <-volameAGV %>% anova_test(dv = VFAt_valerate, wid = EDE, within = Time, between =groups)
get_anova_table(res.aov_to)


res.aov_to <-volameAGV %>% anova_test(dv = VFAt_isovalerate, wid = EDE, within = Time, between =groups)
get_anova_table(res.aov_to)

res.aov_to <-volameAGV %>% anova_test(dv = VFAt_butyrate, wid = EDE, within = Time, between =groups)
get_anova_table(res.aov_to)

```


```{r}
#Post hoc # comparaisons par paires
#volameAGV<-na.omit(volameAGV)

pair<-volameAGV %>% 
pairwise_t_test( VFA_totaux~Lot_x_Time, p.adjust.method = "bonferroni" ) 
data.frame(pair) 

pair_propionate<-volameAGV %>% 
pairwise_t_test( VFAt_propionate~Lot_x_Time, p.adjust.method = "bonferroni" ) 
data.frame(pair_propionate) 

pair

pair_propionate

#Procedure for a significant two-way interaction 
#VFAtoto + VFA propionate

#VFAtoto
# Effect of treatment at each time point
one.way <- volameAGV %>%
  group_by(Time) %>%
  anova_test(dv = VFA_totaux, wid = EDE, between = groups) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
one.way

# Pairwise comparisons between treatment groups
pwc <- volameAGV %>%
  group_by(Time) %>%
  pairwise_t_test(
    VFA_totaux ~ groups, paired = TRUE,
    p.adjust.method = "bonferroni"
    )
pwc

```

Considering the Bonferroni adjusted p-value (p.adj), it can be seen that the simple main effect of treatment was not significant at the time point t3 and t13 (p = 1). It becomes significant at t10.


Pairwise comparisons show that the mean VFAtt score was significantly different between DAm and Mixed at t10 (p = 0.011) and Mixed ocntrol (p = 0.006) 

```{r}

#VFA propionate

# Effect of treatment at each time point
one.way <- volameAGV %>%
  group_by(Time) %>%
  anova_test(dv = VFAt_propionate, wid = EDE, between = groups) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
one.way

# Pairwise comparisons between treatment groups
pwc <- volameAGV %>%
  group_by(Time) %>%
  pairwise_t_test(
    VFAt_propionate ~ groups, paired = TRUE,
    p.adjust.method = "bonferroni"
    )
pwc
```
Considering the Bonferroni adjusted p-value (p.adj), it can be seen that the simple main effect of treatment was not significant at the time point t3 (p = 1). It becomes significant at t13 +10 


Pairwise comparisons show that the mean VFAtt score was significantly different between control and Mixed at t10 (p = 0.014) and 13 : Dam vs Mixed  (p = 0.007) 


MANOVA 

Assumptions and preleminary tests

```{r}

#Multivariate normality
volameAGV %>%
  select(VFA_totaux, VFAt_Acetate,VFAt_butyrate,VFAt_caproate, VFAt_isovalerate,VFAt_isobutyrate, VFAt_valerate,VFAt_propionate) %>%
  mshapiro_test()

#Identify multicollinearity
volameAGV %>% cor_test(VFA_totaux, VFAt_Acetate)

#lineratuty
# Create a scatterplot matrix by group
library(GGally)
results <- volameAGV %>%
  select(VFA_totaux, VFAt_Acetate,VFAt_butyrate,VFAt_caproate, VFAt_isovalerate,VFAt_isobutyrate, VFAt_valerate,VFAt_propionate, Lot_x_Time) %>%
  group_by(Lot_x_Time) %>%
  doo(~ggpairs(.) + theme_bw(), result = "plots")
results
# Show the plots
results$plots

#Check the homogeneity of covariances assumption
box_m(volameAGV[, c("VFA_totaux", "VFAt_Acetate","VFAt_butyrate","VFAt_caproate", "VFAt_isovalerate","VFAt_isobutyrate", "VFAt_valerate","VFAt_propionate")], volameAGV$Lot_x_Time)

#Check the homogneity of variance assumption
volameAGV %>% 
  gather(key = "variable", value = "value", VFA_totaux, VFAt_Acetate,VFAt_butyrate,VFAt_caproate, VFAt_isovalerate,VFAt_isobutyrate, VFAt_valerate,VFAt_propionate) %>%
  group_by(variable) %>%
  levene_test(value ~ Lot_x_Time)
```

```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(car)
library(broom)
# MANOVA test
res.man <- manova(cbind(VFA_totaux, VFAt_Acetate,VFAt_butyrate,VFAt_caproate, VFAt_isovalerate,VFAt_isobutyrate, VFAt_valerate,VFAt_propionate) ~ Time*Lot, data = volameAGV)

summary(res.man,tol=0)

summary.aov(res.man)

```

