---
title: "AQ_volame_AQ1224_clinical_disorder"
output: html_document
date: "2023-01-30"
---

setwd("Z:/DINAMIC/Echanges_nontit/Laurianne_Voland/AQ_Volame/AQ_Volame_Health")

### I- Library and data 
```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(rstatix)
library(readxl)

CD<- read_excel("Health/DOA_disorders.xlsx")
View(CD)
```


```{r}
table(CD$Time,CD$LOT,CD$Respiratoire)
table(CD$Time,CD$LOT,CD$Diarrhees)
```


# CHI
```{r}
stu_data = data.frame(CD$Lot_x_Time, CD$Diarrhees) 
stu_data = table(CD$Lot_x_Time, CD$Diarrhees)
print(chisq.test(stu_data)) 

stu_data_R = data.frame(CD$Lot_x_Time, CD$Respiratoire) 
stu_data_R = table(CD$Lot_x_Time, CD$Respiratoire)
print(chisq.test(stu_data_R)) 
```
Lot * Time  p = 0.003272 


Analysis of binary repeated measures data with R : For outcomes that are binary a linear model is not reasonable since the mean is constrained to the [0; 1]-interval.
Instead we use logistic regression to model the log-odds of an outcome as dependent on covariates

```{r setup, include=FALSE}
#multivariable regression 
hoops=glmer(Diarrhees~Lot*Time+(1|EDE),family=binomial,data=CD)
Anova(hoops, type="III") # Each effect controlled for all others

hoops_respi=glmer(Respiratoire~Lot*Time+(1|EDE),family=binomial,data=CD)
Anova(hoops_respi, type="III") # Each effect controlled for all others

# Because the interaction is significant but close, try a likelihood ratio test
# Fit a model with main effects but no interaction, and compare
restrictedmodel = glmer(Diarrhees ~ Lot+Time + (1 | EDE), family=binomial, data=CD)
anova(restrictedmodel,hoops)
```

NS
